---
title: "Hands-on Exercise 1"
author: "Xu Xinyi"
---

# Getting Started

## Install and launching R packages

The code chunk below uses p_load() of pacman packages to check if tidyverse packages are installed in the computer. If they are, then they will be launched into R.

```{r}
pacman::p_load(tidyverse)
```

## Importing the data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

# Plotting a simple bar chart

```{r}
 

ggplot(data = exam_data,
       aes(x= RACE))+
  geom_bar()
```

Essential Grammatical Elements in ggplot2: data
 
```{r}
ggplot(data=exam_data)
```

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS))
```

Geometric Objects: geom_bar

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar()
```

Geometric Objects: geom_dotplot

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot(dotsize = 0.5)
```
```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot(binwidth=2.5,         
               dotsize = 0.5) +      
  scale_y_continuous(NULL,           
                     breaks = NULL)  
```

Geometric Objects: geom_histogram()

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_histogram()       
```

 Modifying a geometric object by changing geom()
 
```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20,            
                 color="black",      
                 fill="light blue")  
```

