---
title: "Take-home Exercise 03.b"
author: "xu xinyi"
format: html
editor: visual
date-modified: "last-modified"
description: "Prototyping Modules for Visual Analytics Shiny Application"
categories: [tidyverse, jsonlite,tidygraph,igraph,ggraph,shiny,lubridate,DT]
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

# Overview

This prototype module focuses on the Network Graph component of our Shiny application for the VAST Challenge 2025 Mini-Challenge 1. The goal is to visualize the influence and collaborations of Sailor Shift and other Oceanus Folk artists over time using interactive network analytics.

The network graph enables exploration of musical relationships, genre diffusion, and temporal influence patterns. This report documents the prototype development, including data preparation, package evaluation, UI design, visual principles, and tested code.

# Loading R packages

::: panel-tabset
## Description

| Package     | Description                     |
|-------------|---------------------------------|
| `tidyverse` | Data manipulation and cleaning  |
| `jsonlite`  | Reading JSON data               |
| `igraph`    | Network analysis                |
| `ggraph`    | Network visualization           |
| `shiny`     | Web application framework       |
| `lubridate` | Date handling                   |
| `DT`        | Interactive tables for Shiny UI |

## Loading

```{r}
library(jsonlite)
library(tidyverse)
library(tidygraph)
library(ggraph)
library(lubridate)
library(igraph)

# Load JSON graph
kg <- fromJSON("MC1_graph.json")
```
:::

# Extracting the edges and nodes tables

`as_tibble()` of tibble package package is used to extract the nodes and links tibble data frames from kg object into two separate tibble data frames called nodes_tbl and edges_tbl respectively.
```{r}
nodes_tbl <- as_tibble(kg$nodes)
edges_tbl <- as_tibble(kg$links) 
```

# Creating Knowledge Graph

## Mapping from node id to row index
It is important for us to ensures each id from the node list is mapped to the correct row number. This requirement can be achive by using the code chunk below.
```{r}
id_map <- tibble(id = nodes_tbl$id,
                 index = seq_len(
                   nrow(nodes_tbl)))
```

## Map source and target IDs to row indices
We will map the source and the target IDs to row indices by using the code chunk below.
```{r}
edges_tbl <- edges_tbl %>%
  left_join(id_map, by = c("source" = "id")) %>%
  rename(from = index) %>%
  left_join(id_map, by = c("target" = "id")) %>%
  rename(to = index)
```

## Filter out any unmatched (invalid) edges

```{r}
edges_tbl <- edges_tbl %>%
  filter(!is.na(from), !is.na(to))
```

## Creating tidygraph
`tbl_graph()` is used to create tidygraph’s graph object by using the code chunk below.
```{r}
graph <- tbl_graph(nodes = nodes_tbl, 
                   edges = edges_tbl, 
                   directed = kg$directed)
```

You might want to confirm the output object is indeed in tidygraph format by using the code chunk below.

```{r}
class(graph)
```

# Subgraph Extraction by Artist and Date

Extract a subgraph centered around a specific artist (e.g., Sailor Shift) and filter by a date window (e.g., songs released before or after a certain year).

```{r}
sailor_node <- nodes_tbl %>% filter(name == "Sailor Shift") %>% pull(id)
sub_nodes <- neighborhood(graph, order = 2, nodes = sailor_node) %>% unlist()
subgraph <- induced_subgraph(graph, sub_nodes)
```

# Centrality Measures: PageRank and Betweenness

```{r}
graph <- graph %>%
  mutate(
    pagerank = centrality_pagerank(),
    betweenness = centrality_betweenness()
  )
```

```{r}
graph %>%
  as_tibble() %>%
  arrange(desc(pagerank)) %>%
  select(name, genre, pagerank, betweenness) %>%
  head(10)
```

 
